Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-10-08 paper.tex
almost ready to go

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a$ and $\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$. 
While most hyper-Raman experiments are performed using degenerate input pulses ($\omega_a = \omega_b$), it is possible to perform the technique with non-degenerate input pulses ($\omega_a \neq \omega_b$). \cite{Denisov1986, Kozich2007}
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through a two-photon inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been observed in several CMDS studies (sometimes under the moniker of SIVE).\cite{Zilian1994, RN350, RN351, RN352, RN353, Chen1998, RN362, RN418, Hanninen2018, Wang2021, Bonn2024, McDonnell2024}

@@

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares them with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require macroscopic non-centrosymmetry under the electric dipole approximation.\cite{RN480}
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to four-wave mixing (FWM) cascades.\cite{RN515, RN243, Cho2000_Cascade}
A coherent six-wave mixing analogue of hyper-Raman scattering, coherent anti-Stokes hyper-Raman spectroscopy (CAHRS), has been proposed and explored theoretically by several authors; however, it is also weak and highly susceptible to FWM cascades.\cite{Taran1977, Berger1978, Bjarnason1980, Cho1997, Cho1998}
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave and six-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just a single vibrational mode and electronic state (``evx'' coupling).\cite{RN445, RN335} 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as two-color experiments (with $\omega_2=\omega_3$);\cite{Cho2001} vibrational sum frequency generation (vSFG) or difference frequency generation (vDFG) setups can be trivially reconfigured to perform HSFG or HDFG measurements to investigate ev coupling in bulk systems.

@@

This paper investigates the microscopic parameters that control HDFG output. \cite{RN352, Bonn2024, McDonnell2024}
HSFG will not be discussed, but the application of our treatment to HSFG is straightforward and the results are similar.
We note that HDFG emission can be phase-matched in media with normal dispersion while HSFG cannot.\cite{RN279, RN278} 
In \autoref{steadystate}, we first identify the vibrational selection rules. 
The selection rules of electronically-resonant HDFG are then identified through an expansion of electronic dipole moments following Chung and Ziegler.\cite{Ziegler1988}
HDFG is found to be allowed in any harmonic system that has infrared active vibrations.
A simple harmonic oscillator model system is used to simulate HDFG spectra and illustrate how the excited state potential energy surface affects the hyper-Raman excitation spectrum.
After developing selection rules, some quantitative aspects of HDFG are discussed in \autoref{quant}.

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
Coherent spectroscopies are often described using the density matrix formalism in the interaction picture, where an n$^{\text{th}}$ order density matrix $\rho^{(n)}$ is propagated using the Liouville-von Neumann equation:
\begin{equation}\label{LVNE}
	\dot{\rho}^{(n)} = \frac{1}{i \hbar} [H + V(t), \rho^{(n-1)}] + \dot{\rho}^{(n)}_R,
\end{equation}
where $H$ is the (static) system Hamiltonian, $V(t)$ is the field-matter interaction and $\rho_R$ is the relaxation density matrix. \cite{RN455}
In systems where interactions with surroundings are relevant, expressions for $\dot{\rho}_R$ may be quite complicated due to intricate phenomena such as system-bath interactions and solvent-solute coupling, among others. \cite{Sung2001}
In the limit where input field pulse widths are much longer than electronic dephasing times and the individual molecule limit is approached, it becomes acceptable to write a time-independent relaxation matrix $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$, i.e. exponential relaxation, where $\Gamma$ is the system's dephasing rate. \cite{RN455, Sue1986, Sung2001}
While exponential relaxation is usually an inappropriate description of most physical systems involving solvent-solute interactions,\cite{Sue1986, Yan1988, Li1994, Myers1997} our goal is to introduce simple, quantitative models to understand HDFG response.
Thus, we write $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$.
The exponential relaxation limit has proven useful in investigating the properties of 2D-EV, 2D-VE, and doubly resonant SFG using pulses much longer than electronic dephasing times. \cite{Raschke2002, Gaynor2017}

It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.\cite{Cho2008}
While we have only isolated one specific HDFG resonance and time-ordering scheme (\autoref{fig:comparisonwmel}), there are several other pathways which can provide HDFG output. \cite{RN352}
If $\omega_2$ becomes resonant and $\omega_1$ becomes non-resonant, where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the pathways presented in \autoref{fig:sivewmel2} will interfere, which can amplify or obscure coherent output.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=\textwidth]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$, described in \autoref{Albrecht}.
		In all cases, $\omega_2 > \omega_1$.
		The virtual states fluctuate in energy dependent upon the energy of the input fields.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ is detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ interacts first and is resonant, and $\omega_2 + \omega_3$ is nonresonant, yields two contributed pathways: \autoref{fig:sivewmel2} (a) and (b).
These pathways possess identical phase and transition moments, which yields total constructive interference, assuming similar non-resonant dispersion from each pathway.
Conversely, in the case where only $\omega_2$ is vibrationally resonant, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned from an electronic state. \cite{McDonnell2024}
Note that in the case where $\omega_1$ interacts first but $\omega_2$ is resonant, the nonlinear output generated nearly vanishes as the first resonant pulse arrives after a non-resonant pulse.

When $\omega_2 + \omega_3$ becomes electronically resonant, the interference changes for the \autoref{fig:sivewmel2}c-f pathways.
Note that in this regime, only pathways (c,e,f) possess identical transition moments.
As a result, pathways (c,d) survive and can provide output when $\omega_2+\omega_3$ is resonant with an electronic state. 
These results provide a simple rule of thumb for collecting HDFG spectra: use \autoref{fig:sivewmel2}a,b pathways to measure spectra when significantly detuned from an electronic resonance, and any of these vibrationally resonant pathway can measure spectra when electronically resonant.
However, since electronic states commonly have short dephasing times and significant inhomogeneous broadening,\cite{Dong2015, Lewis2015} it is likely the \autoref{fig:sivewmel2}f pathway will become triply resonant and possibly contribute greater than the other pathways. 
Such instances are system dependent. 
Discussion will be isolated to the \autoref{fig:sivewmel2}a pathway in the sections that follow.

@@

\subsection{Vibronic Selection Rules}\label{Albrecht}
While the Placzek treatment provides the general source of HDFG output, it does not predict the behavior of $\beta_{ijk}$ as $\omega_2 + \omega_3$ is changed.
We now explicitly consider electronic resonances through the $A,B,C$ decomposition of $\beta$ introduced by Chung and Ziegler, analogous to Albrecht's treatment of Raman spectroscopy (i.e., sum-over-states formalism).\cite{Albrecht1961, Ziegler1988} 
To employ this formalism, we write the states in a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(\textbf{Q})) \otimes \ket{b}$ for electronic states $\{|a(\textbf{Q}))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation), where \textbf{Q}, the system normal modes, is the adiabatic parameter. \cite{BornOppenheimer, Tang1970}
We will henceforth suppress the dependence of the electronic states on \textbf{Q} for simplicity.
It is worthwhile noting that the time-dependent Raman formalism developed by Heller \textit{et al.} can be applied to hyper-Raman scattering. \cite{Lee1979, Silverstein2012}
However, since our goal is to expose the Quantum Mechanical aspects of HDFG spectroscopy, we choose to employ the typical sum-over-states formalism.

For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments; $\vec{\mu}$ is reserved for transitions on the ground electronic state. \cite{Tang1970}
Following the approach which gave \autoref{sivegamma} and using the time ordering and state labeling of \autoref{fig:sivewmel2}a, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{4D \hbar^3} \sum_{m,n,e,\tilde{v}'} \frac{
		R_{i}^{gv, e\tilde{v}'} 
		R_{j}^{e\tilde{v}',mn} 
		R_{k}^{mn,g0} 
		R_{l}^{g0,gv} 
	}{	\Delta_{e\tilde{v}', gv}
		\Delta_{mn, gv}
		\Delta_{g0,gv}
	},
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.
To ease readability, we use $\ket{a}$ to refer to ground state vibrational modes and $\ket{\tilde{b}}$ to refer to excited state vibrational modes.
Using our definition of vibronic states, we can write, for example,
$R_{i}^{gv,e\tilde{v}'} = \mel{v}{M_i^{ge}}{\tilde{v}'}$, where $\vec{M}^{ab} = (a|\vec{R}|b)$.\cite{Ziegler1974}
Expanding $\vec{M}^{ij}$ to $\order{Q}$ about the equilibrium point of the ground state potential surface (taken to be $Q = 0$) as
$\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z} Q_z$
yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Warshel1977, Ziegler1988} so that
\begin{equation}\label{gammaabc}
		\gamma_{ijkl} \sim \left(A_{ijk} + B_{ijk} + C_{ijk}\right) \frac{\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}.
\end{equation}
The $A$ term contains the static ($\order{Q^0}$) transitions (i.e., Condon approximation),\cite{Condon1928} the $B$ term depends upon $\order{Q}$ transitions (Herzberg-Teller contributions),\cite{HerzbergTeller1933} and the $C$ term depends on $\order{Q_i Q_j}$ transitions. 
The $C$ term is suppressed in the following discussion as it depends on one and two photon forbidden transitions, making its contribution to $\gamma_{ijkl}$ roughly two orders of magnitude less than $A_{ijk}$. \cite{Ziegler1988, Neddersen1989, Bonang1992}
Note that some reports obtain $A, B$ coefficients through a Herzberg-Teller expansion of the electronic states to expose vibronic couplings via $\partial H / \partial Q$.\cite{HerzbergTeller1933, Petrov1985, Neddersen1989, Baranov1990}
This approach is not used here as the expansion of $\vec{M}^{ij}$ in normal mode coordinates provides sufficient physical insight into HDFG selection rules. 
Duschinsky rotation effects\cite{Duschinsky1937} are ignored.

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$,\cite{Milojevich2013} are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,\tilde{v}'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | \tilde{v}' \rangle
		 \langle \tilde{v}' | 0 \rangle 
		 \frac{1}{\Delta_{e\tilde{v}', gv}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | \tilde{v}' \rangle \mel{\tilde{v}'}{Q_z}{0} \frac{1}{\Delta_{e\tilde{v}', gv}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{\tilde{v}'} 
			\langle \tilde{v}' | 0 \rangle 
			\frac{1}{\Delta_{e\tilde{v}', gv}}.
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
Note that it is common in the hyper-Raman literature to write $\Delta_{e\tilde{v}', gv} \Delta_{mn, gv} = \Delta_{e\tilde{v}', gv}$.
Here, $\langle \tilde{a} | b \rangle$ and $\mel{\tilde{a}}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively.
At this point, the expressions are valid for both electronically resonant and non-resonant cases.

To understand the vibronic structure of an excited state $|e)$, we restrict consideration to electronic transitions between $|g)$ and $|e)$.  
Note that HDFG experiments selectively excite ground state vibrational modes through $\omega_1$; we will assume the IR beam has excited the fundamental of an IR-active mode, $\ket{g,1}$, so that hyper-Raman scattering is between the ground state and $\ket{g,1}$. 
By working only in terms of this single normal mode (i.e., removing the sum over $z$ and taking $v=1$), 
the $A$ and $B$ terms are rewritten as 
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
				A_{ijk} = \frac{1}{\hbar}\sum_{\tilde{v}'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | \tilde{v}' \rangle
				\langle \tilde{v}' | 0 \rangle 
				\frac{1}{\Delta_{e\tilde{v}',g1}}
		\end{equation}
		\begin{equation}
				B_{1_{ijk}} = \frac{1}{\hbar} \sum_{\tilde{v}'} M^{ge}_{0,i}
				\frac{\partial \Lambda^{eg}_{jk}}{\partial Q} \langle 1 | \tilde{v}' \rangle \mel{\tilde{v}'}{Q}{0} 
				\frac{1}{\Delta_{e\tilde{v}', g1}}
		\end{equation}
		\begin{equation}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{\tilde{v}'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{\tilde{v}'} 
				\langle \tilde{v}' | 0 \rangle 
				\frac{1}{\Delta_{e\tilde{v}',g1}}
		\end{equation}
	\end{subequations}
where $\Lambda^{ab}_{ij} = 1/\hbar \sum_m M_i^{am}M_j^{mb} $ is the electronic two photon transition moment between $|a)$ and $|b)$.

>>>>

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a$ and $\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$. 
While most hyper-Raman experiments are performed using degenerate input pulses ($\omega_a = \omega_b$),\cite{RN515} it is possible to perform the technique with non-degenerate input pulses ($\omega_a \neq \omega_b$). \cite{Denisov1986, Kozich2007}
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through a two-photon inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been observed in several CMDS studies (sometimes under the moniker of SIVE, or singly vibrationall enhanced spectroscopy).\cite{Zilian1994, RN350, RN351, RN352, RN353, Chen1998, RN362, RN418, Hanninen2018, Wang2021, Bonn2024, McDonnell2024}

@@

The HDFG and HSFG processes and related spectroscopies that investigate vibrational-electronic coupling are diagrammed in \autoref{fig:comparisonwmel} using Wave Mixing Energy Level (WMEL) diagrams. \cite{RN286}
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require macroscopic non-centrosymmetry under the electric dipole approximation.\cite{RN480}
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to four-wave mixing (FWM) cascades.\cite{RN515, RN243, Cho2000_Cascade}
A coherent six-wave mixing analogue of hyper-Raman scattering, coherent anti-Stokes hyper-Raman spectroscopy (CAHRS), has been proposed and explored theoretically by several authors; however, it is also weak and susceptible to FWM cascades.\cite{Taran1977, Berger1978, Bjarnason1980, Cho1997, Cho1998}
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave and six-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just a single vibrational mode and electronic state (``evx'' coupling).\cite{RN445, RN335} 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as two-color experiments (with $\omega_2=\omega_3$);\cite{Cho2001} vibrational sum frequency generation (vSFG) or difference frequency generation (vDFG) setups can be trivially reconfigured to perform HSFG or HDFG measurements to investigate ev coupling in bulk systems.

@@

This paper investigates the microscopic parameters that control HDFG output. \cite{RN352, Bonn2024, McDonnell2024}
HSFG will not be discussed, but the application of our treatment to HSFG is straightforward and the results are similar.
We note that HDFG emission can be phase-matched in media with normal dispersion while HSFG cannot.\cite{RN279, RN278} 
This paper is organized as follows.
In \autoref{steadystate}, we first identify the HDFG vibrational selection rules. 
The electronically-resonant HDFG selection rules are then identified through an expansion of electronic dipole moments following Chung and Ziegler.\cite{Ziegler1988}
HDFG is found to be allowed in any harmonic system that has infrared active vibrations.
A simple harmonic oscillator model system is used to simulate HDFG spectra and illustrate how the excited state potential energy surface affects the hyper-Raman excitation spectrum.
After developing selection rules, some quantitative aspects of HDFG are discussed in \autoref{quant}.

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
Coherent spectroscopies are often described using the density matrix formalism in the interaction picture, where an n$^{\text{th}}$ order density matrix $\rho^{(n)}$ is propagated using the Liouville-von Neumann equation:
\begin{equation}\label{LVNE}
	\dot{\rho}^{(n)} = \frac{1}{i \hbar} [H + V(t), \rho^{(n-1)}] + \dot{\rho}^{(n)}_R,
\end{equation}
where $H$ is the (static) system Hamiltonian, $V(t)$ is the field-matter interaction and $\rho_R$ is the relaxation density matrix. \cite{RN455}
In systems where interactions with surroundings are relevant, expressions for $\dot{\rho}_R$ may be quite complicated due to intricate phenomena such as system-bath interactions and solvent-solute coupling, among others. \cite{Sung2001}
In the limit where input field pulse widths are much longer than electronic dephasing times and the individual molecule limit is approached, it becomes acceptable to write a time-independent relaxation matrix $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$, i.e., exponential relaxation, where $\Gamma$ is the system's dephasing rate. \cite{RN455, Sue1986, Sung2001}
While exponential relaxation is usually an inappropriate description of most physical systems involving solvent-solute interactions,\cite{Sue1986, Yan1988, Li1994, Myers1997} our goal is to introduce simple, quantitative models to understand HDFG response.
Thus, we write $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$.
The exponential relaxation limit has proven useful in investigating the properties of 2D-EV, 2D-VE, and doubly resonant SFG using pulses much longer than electronic dephasing times. \cite{Raschke2002, Gaynor2017}

It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.\cite{Cho2008}
While we have only isolated one specific HDFG resonance and time-ordering scheme (\autoref{fig:comparisonwmel}), there are several other pathways which can provide HDFG output. \cite{RN352}
If $\omega_2$ becomes resonant and $\omega_1$ becomes non-resonant, where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}c-f).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the pathways presented in \autoref{fig:sivewmel2} will interfere, which can amplify or decrease coherent output.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=\textwidth]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		In (a,b), $\omega_1$ is resonant with a vibrational mode, but in (c-f), $\omega_2$ is resonant with a vibrational mode.
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$, described in \autoref{Albrecht}.
		In all cases, $\omega_2 > \omega_1$.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ is detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ interacts first and is resonant, and $\omega_2 + \omega_3$ is nonresonant, yields two contributed pathways: \autoref{fig:sivewmel2} (a) and (b).
These pathways possess identical phase and transition moments, which yields total constructive interference, assuming similar non-resonant dispersion from each pathway.
Conversely, in the case where only $\omega_2$ is resonant, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned from an electronic state. \cite{McDonnell2024}
Note that in the case where $\omega_1$ interacts first but only $\omega_2$ is resonant, the nonlinear output generated nearly vanishes as the first resonant pulse arrives after a non-resonant pulse.

When $\omega_2 + \omega_3$ becomes electronically resonant, the interference changes for the \autoref{fig:sivewmel2}c-f pathways.
Note that in this regime, only pathways (c,e,f) possess identical transition moments.
As a result, pathways (c,d) survive and can provide output when $\omega_2+\omega_3$ is resonant with an electronic state. 
These results provide a simple rule of thumb for collecting HDFG spectra: use \autoref{fig:sivewmel2}a,b pathways to measure spectra when significantly detuned from an electronic resonance, and any of these vibrationally resonant pathways can measure spectra when electronically resonant.
However, since electronic states commonly have short dephasing times and significant inhomogeneous broadening,\cite{Dong2015, Lewis2015} it is likely the \autoref{fig:sivewmel2}f pathway will become triply resonant and possibly contribute greater than the other pathways. 
Such instances are system dependent. 
Discussion will be isolated to the \autoref{fig:sivewmel2}a pathway in the sections that follow.

@@

\subsection{Vibronic Selection Rules}\label{Albrecht}
While the Placzek treatment provides the source of vibrationally HDFG output, it does not predict the behavior of $\beta_{ijk}$ as $\omega_2 + \omega_3$ is changed.
We now explicitly consider electronic resonances through the $A,B,C$ decomposition of $\beta$ introduced by Chung and Ziegler, analogous to Albrecht's treatment of Raman spectroscopy (i.e., sum-over-states formalism).\cite{Albrecht1961, Ziegler1988} 
To employ this formalism, we write the states in a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(\textbf{Q})) \otimes \ket{b}$ for electronic states $\{|a(\textbf{Q}))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation), where \textbf{Q}, the system normal modes, is the adiabatic parameter. \cite{BornOppenheimer, Tang1970}
We will henceforth suppress the dependence of the electronic states on \textbf{Q} for simplicity.
It is worthwhile noting that the time-dependent Raman formalism developed by Heller \textit{et al.} can be applied to hyper-Raman scattering. \cite{Lee1979, Shoute2005, Silverstein2012}
However, since our goal is to expose the Quantum Mechanical aspects of HDFG spectroscopy, we choose to employ the typical sum-over-states formalism.

For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments; $\vec{\mu}$ is reserved for transitions on the ground electronic state. \cite{Tang1970}
Following the approach which gave \autoref{sivegamma} and using the time ordering and state labeling of \autoref{fig:sivewmel2}a, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{4D \hbar^3} \sum_{m,n,e,\tilde{v}'} \frac{
		R_{i}^{gv, e\tilde{v}'} 
		R_{j}^{e\tilde{v}',mn} 
		R_{k}^{mn,g0} 
		R_{l}^{g0,gv} 
	}{	\Delta_{e\tilde{v}', gv}
		\Delta_{mn, gv}
		\Delta_{g0,gv}
	},
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.
To ease readability, we use $\ket{a}$ to refer to ground state vibrational modes and $|\tilde{b}\rangle$ to refer to excited state vibrational modes.
Using our definition of vibronic states, we can write, for example,
$R_{i}^{gv,e\tilde{v}'} = \mel{v}{M_i^{ge}}{\tilde{v}'}$, where $\vec{M}^{ab} = (a|\vec{R}|b)$.\cite{Ziegler1974}
Expanding $\vec{M}^{ij}$ to $\order{Q}$ about the equilibrium point of the ground state potential surface (taken to be $Q = 0$) as
$\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z} Q_z$
yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Warshel1977, Ziegler1988} so that
\begin{equation}\label{gammaabc}
		\gamma_{ijkl} \sim \left(A_{ijk} + B_{ijk} + C_{ijk}\right) \frac{\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}.
\end{equation}
The $A$ term contains the static ($\order{Q^0}$) transitions (i.e., Condon approximation),\cite{Condon1928} the $B$ term depends upon $\order{Q}$ transitions (Herzberg-Teller contributions),\cite{HerzbergTeller1933} and the $C$ term depends on $\order{Q_i Q_j}$ transitions. 
The $C$ term is suppressed in the following discussion as it depends on one and two photon forbidden transitions, making its contribution to $\gamma_{ijkl}$ roughly two orders of magnitude less than $A_{ijk}$. \cite{Ziegler1988, Neddersen1989, Bonang1992}
Note that some reports obtain $A, B$ coefficients through a Herzberg-Teller expansion of the electronic states to expose vibronic couplings via $\partial H / \partial Q$.\cite{HerzbergTeller1933, Petrov1985, Neddersen1989, Baranov1990}
This approach is not used here as the expansion of $\vec{M}^{ij}$ in normal mode coordinates provides sufficient physical insight into HDFG selection rules. 
Duschinsky rotation effects\cite{Duschinsky1937} are ignored.

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$,\cite{Milojevich2013} are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,\tilde{v}'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | \tilde{v}' \rangle
		 \langle \tilde{v}' | 0 \rangle 
		 \frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | \tilde{v}' \rangle \mel{\tilde{v}'}{Q_z}{0} \frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{\tilde{v}'} 
			\langle \tilde{v}' | 0 \rangle 
			\frac{1}{\Delta_{e\tilde{v}', gv} \Delta_{m, g}}.
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
Note that to allow closure over $\ket{n}$, we write $\Delta_{mn, gv} \approx \Delta_{m, g}$, which is allowed since the system is assumed to be sufficently detuned from $\ket{n}$. \cite{Silverstein2012}
Here, $\langle \tilde{a} | b \rangle$ and $\mel{\tilde{a}}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively.
At this point, the expressions are valid for both electronically resonant and non-resonant cases.

To understand the vibronic structure of an excited state $|e)$, we restrict consideration to electronic transitions between $|g)$ and $|e)$.  
Note that HDFG experiments selectively excite ground state vibrational modes through $\omega_1$; we will assume the IR beam has excited the fundamental of an IR-active mode, $\ket{g,1}$, so that the hyper-Raman scattering is between the ground state and $\ket{g,1}$. 
By working only in terms of this single normal mode (i.e., removing the sum over $z$ and taking $v=1$), 
the $A$ and $B$ terms are rewritten as 
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
				A_{ijk} = \frac{1}{\hbar}\sum_{\tilde{v}'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | \tilde{v}' \rangle
				\langle \tilde{v}' | 0 \rangle 
				\frac{1}{ \Delta_{e\tilde{v}',g1}}
		\end{equation}
		\begin{equation}
				B_{1_{ijk}} = \frac{1}{\hbar} \sum_{\tilde{v}'} M^{ge}_{0,i}
				\frac{\partial \Lambda^{eg}_{jk}}{\partial Q} \langle 1 | \tilde{v}' \rangle \mel{\tilde{v}'}{Q}{0} 
				\frac{1}{\Delta_{e\tilde{v}', g1}}
		\end{equation}
		\begin{equation}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{\tilde{v}'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{\tilde{v}'} 
				\langle \tilde{v}' | 0 \rangle 
				\frac{1}{\Delta_{e\tilde{v}',g1}}
		\end{equation}
	\end{subequations}
where $\Lambda^{ab}_{ij} = 1/\hbar \sum_m M_i^{am}M_j^{mb} \Delta^{-1}_{m, b} $ is the electronic two photon transition moment between $|a)$ and $|b)$.
